cleaning script for NationalOutbreakPublicDataTool.xlsx

### 2023-02-23 code/processing_code/processingfile_v1.qmd
3

---
title: "An example cleaning script with code pulled in"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---

# Processing script with code pulled in

This is essentially the same as the other Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script using the code chunk labels.

# Setup

This needs to run to load the R script so we can include the code chunks below.

Note that you can pull in more than one R script if you want.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('processingcode.R')
```

Load needed packages. Instead of having R commands here, this just pulls in the code from the R script we loaded above and labelled.

```{r, packages, message = FALSE, warning = FALSE}
library(readxl) #for loading excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(dlookr)
```

@@

```{r, loaddata}
data_location<-here("data","raw_data","NationalOutbreakPublicDataTool.xlsx")
rawdata2=read_excel(data_location)

```

# Explore data

@@

```{r, exploredata, fig.height=8, fig.width=8}
glimpse(rawdata2)
summary(rawdata2)
head(rawdata2)
skim(rawdata2)
plot_na_pareto(rawdata2)
```

Looks like we have the following data:

* height (seems like it's in centimeters) 
* weight (seems to be in kilogram)
* Sex

@@

First, there is an entry for height which says "sixty" instead of a number. 
Does that mean it should be a numeric 60? It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
Since we don't know how to fix this, we might decide to remove the person. This "sixty" entry also turned all Height entries into characters instead of numeric. That conversion to character also means that our summary function isn't very meaningful. So let's fix that first.

```{r, cleandata1,fig.height=8, fig.width=8}
d2=rawdata2%>%
  filter(`Primary Mode`=="Food")
plot_na_pareto(d2)
```

@@

```{r, cleandata2}
```

@@

Note that removing anyone who had "faulty" or missing data is one approach. It's often not the best. Based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

```{r, cleandata3}
```

@@

```{r, cleandata4}
```

@@

```{r, cleandata5}
```

@@

All done, data is clean now. Time to save.
Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

@@

```{r, savedata}
```

>>>>

---
title: "cleaning script for foodborne_outbreaks.csv"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---

# Setup
```{r}
library(readr) #for loading csv files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(dlookr)
```

@@

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw_data","foodborne_outbreaks.csv")
rawdata1 <- read_csv(data_location)
```

# Check data

@@

```{r}
str(rawdata1)
summary(rawdata1)
head(rawdata1)
skim(rawdata1)
plot_na_pareto(rawdata1)
```

Looks like we have the following predictor variables:

* Year, Month (date)
* Sate, Location (geographic information)
* Food, Ingredient
*Species, Serotype/Genotype
*Status

Following variables are outcomes:
*illnesses
*Hospitalizations
*Fatalities

@@

First, we will remove the variable "ingredient" and "serotype" since over 80% of data is missing. 
Does that mean it should be a numeric 60? It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
Since we don't know how to fix this, we might decide to remove the person. This "sixty" entry also turned all Height entries into characters instead of numeric. That conversion to character also means that our summary function isn't very meaningful. So let's fix that first.

```{r}
d1=
```

@@

```{r}
d2 <- d1 %>% dplyr::filter(Height != 6)
skimr::skim(d2)
```

@@

```{r}
d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
skimr::skim(d3)
```

@@

```{r}
d3$Sex <- as.factor(d3$Sex)  
skimr::skim(d3)
```

@@

```{r}
d4 <- d3 %>% dplyr::filter(Sex != "NA") %>% droplevels()
skimr::skim(d4)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- d4
```

@@

@@

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2023-02-23 code/processing_code/processingfile_v2.qmd
3

---
title: "An example cleaning script with code pulled in"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---

# Processing script with code pulled in

This is essentially the same as the other Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script using the code chunk labels.

@@

This needs to run to load the R script so we can include the code chunks below.

Note that you can pull in more than one R script if you want.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('processingcode.R')
```

Load needed packages. Instead of having R commands here, this just pulls in the code from the R script we loaded above and labelled.

```{r, packages, message = FALSE, warning = FALSE}
library(readxl) #for loading excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(dlookr)
```

@@

```{r, loaddata}
data_location<-here("data","raw_data","NationalOutbreakPublicDataTool.xlsx")
rawdata2=read_excel(data_location)

```

@@

```{r, exploredata, fig.height=8, fig.width=8}
glimpse(rawdata2)
summary(rawdata2)
head(rawdata2)
skim(rawdata2)
plot_na_pareto(rawdata2)
```

@@

```{r, cleandata1,fig.height=8, fig.width=8}
d2=rawdata2%>%
  filter(`Primary Mode`=="Food")
plot_na_pareto(d2)
```

>>>>

---
title: "cleaning script for NationalOutbreakPublicDataTool.xlsx"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---

@@

```{r, include=FALSE, cache=FALSE}
library(readr) #for loading csv files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

```{r, packages, message = FALSE, warning = FALSE}
```

@@

```{r, loaddata}
data_location
```

@@

```{r, exploredata}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

@@

```{r, cleandata1}
```
